---
title: "ANOVA table"
author: "Kaitlyn O'Shea"
date: "2/26/2020"
output: html_document
---

## Proof of Concept for conducting ANOVA across selected clinical variable and output significant results

```{r}
library()
```

```{r}
# use schmidt for example
load("~/Documents/Thesis/mgm_eda/datasets/GSE74385_Schmidt_M_KO.Rda")
```

```{r}
# extract gene names
a <- as.list(colnames(Schmidt_M[["expression_data"]])[-1])
```


```{r}
## baseline test extraction
summary(aov(PTTG1 ~ grade, data = Schmidt_M[["data"]]))[[1]]$'Pr(>F)'[1]
```

### Testing
This chunk helped me to understand the behavior of the dashes and how R natively fixes the column call when directly referencing the data set but not when using variables in the aov statement
```{r}
multi_anova2 <- function(gene, clinical_variable, dataset) {
                                         
    aov <- summary(aov(data = dataset,                           # aov, call the dataset
                                                            formula = (gene         # formula to take gene input
                                                                                    ~
                                                                                    clinical_variable # over clinical variables
                                                                                    )
                                                                              )  
                                                            
                                                        )
                                      print(aov)                                                # print aov 
}

multi_anova2(Schmidt_M[["data"]]$`WI2-2373I1.2`, Schmidt_M[["data"]]$grade, Schmidt_M[["data"]])

```


## This chunk was used to prove that the anova would execute over all gene names in the dataset
```{r}
# for (i in c(1:length(a))) {
#              aov(data = Schmidt_M[["data"]],                           # aov, call the dataset
#                 formula = as.formula(paste0("`",(a[i]),"`",         # formula to take gene input
#                                             "~",
#                                             "grade" # over clinical variables
#                                     )
#                 )
#                 )
#     if (i%%100==0){print(i)}
# }
```

## Formula for functional (app) use

```{r}
multi_anova <- function(gene, clinical_variable, dataset) {
                                          aov <- summary(aov(data = dataset,                           # aov, call the dataset
                                                            # need the quoted backhashes around gene to read "irregular" column/gene names
                                                            formula = as.formula(paste0("`",(gene),"`",  
                                                                                    "~",
                                                                                    clinical_variable # over clinical variables
                                                                                    )
                                                                              )  
                                                            )
                                                        )
                                           anova_pvalue <- data.frame(gene_name = as.character(gene), # gene name
                                                                      `p adj` = aov[[1]]$'Pr(>F)'[1]  # extracted p-value
                                                                      ) 
                                          anova_pvalue
}

multi_anova("PTTG1", "grade", Schmidt_M[["data"]])
multi_anova("WI2-2373I1.2", "grade", Schmidt_M[["data"]])
```

```{r}
X <- multi_anova("PTTG1", "grade", Schmidt_M[["data"]])
str(X)
```


```{r}
genes <- c("EGFR", "NF1", "TP53", "A2M", "PTTG1", "NF2", "PTEN")
```


```{r}
out <- invisible(lapply(genes, multi_anova, clinical_variable = "grade", dataset = Schmidt_M[["data"]]))
do.call(rbind,out)
```

## Putting it all together
```{r}
multi_anova <- function(gene, clinical_variable, dataset) {
                                          aov <- summary(aov(data = dataset,                           # aov, call the dataset
                                                            # need the quoted backhashes around gene to read "irregular" column/gene names
                                                            formula = as.formula(paste0("`",(gene),"`",  
                                                                                    "~",
                                                                                    clinical_variable # over clinical variables
                                                                                    )
                                                                              )  
                                                            )
                                                        )
                                           anova_pvalue <- data.frame(gene_name = as.character(gene), # gene name
                                                                      `p adj` = aov[[1]]$'Pr(>F)'[1]  # extracted p-value
                                                                      ) 
                                          anova_pvalue
}

multianova_out <- invisible(lapply(gene_user_input, multi_anova, clinical_variable = "grade", dataset = Schmidt_M[["data"]]))
multianova_table <- do.call(rbind, multianova_out)
```

