---
title: "GSE85133"
author: "Kaitlyn O'Shea"
output: html_document
---

**THIS SCRIPT HAS BEEN ARCHIVED DUE TO THE AVAILABILITY OF TPM DATA FORMAT ONLY, LIMITING THE ABILITY TO LOOK FOR RELATIONSHIPS ACROSS SAMPLES DUE TO WITHIN SAME NORMALIZATION - CAN REVISIT IF COUNT MATRIX BECOMES PUBLICALLY AVAILABLE**

```{r setup}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE)
```

```{r install_packages, eval = FALSE}
## require BioManager
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("GEOquery") 
```

```{r libraries, message=FALSE, warning=FALSE}
## libraries needed
library("tidyverse")
library("GEOquery")
library("here")
```

## Get GSE85133 from GEO 
```{r Geo_download, message=FALSE, warning=FALSE}
## Download dataset from GEO, save downloaded files in the same folder as this .Rmd file.
dir.create(file.path("raw","gse85133"), recursive = TRUE)
gse85133 <- getGEO("gse85133",  # sample accession 
                   destdir = file.path("raw","gse85133"),
                   GSEMatrix = TRUE,
                   AnnotGPL = TRUE)
```

## Clinical Data from Dataset
```{r ClinicalDataExtract}
## index into data list and extract clinical data
gse85133_cd <- gse85133[["GSE85133_series_matrix.txt.gz"]] %>%   # index into list 
                  phenoData() %>%                                   # select phenoData
                  pData() %>%                                       # apply pData
                  as_tibble() %>%                                   # tidy friendly
                  select("geo_accession" = geo_accession,
                         "sample_id" = `title`,
                         age = `age:ch1`,
                         sex = `gender:ch1`,
                         radiation = `radiated:ch1`,
                         mutant_type = `mutation group:ch1`,
                         grade = `who grade:ch1`,
                         histology = `histology:ch1`,
                         tissue = `tissue:ch1`
                         ) 

gse85133_pheno <- gse85133_cd %>%         # rename and adjust variables
                      mutate(grade = factor(grade,
                                            levels = c("1", "2", "3"),
                                            labels = c("I", "II", "III")
                                            ),
                             sex = tolower(sex),
                             age = as.numeric(age),
                             radiation = factor(radiation,
                                                   levels = c("Y","N", "U"),
                                                   labels = c("yes", "no", "unknown"))

                    ) %>%
                      mutate_if(is.character, as.factor)
```

## rnaseq Data from supplemental GEO files
```{r}
### downlod and and read in supplemental files


## download fpkm_tracking files from GEO sesies supplement
getGEOSuppFiles("GSE85133", makeDirectory = TRUE, fetch_files = TRUE, baseDir = here("datasets_scripts/rnaseq datasets/raw"))

## uncompress the download 
untar(here("datasets_scripts/rnaseq datasets/raw/GSE85133/GSE85133_RAW.tar"), exdir = here("datasets_scripts/rnaseq datasets/raw/GSE85133"))

## unzip all files in _raw directory
zipped <- list.files(path = here("datasets_scripts/rnaseq datasets/raw/GSE85133"), pattern = "*.gz", full.names = TRUE)
sapply(X = zipped, FUN = gunzip)

## read in all unziped files
cuffout <- list.files(path = here("datasets_scripts/rnaseq datasets/raw/GSE85133"), pattern = "*.fpkm_tracking$")
cuff.files <- paste0(here("datasets_scripts/rnaseq datasets/raw/GSE85133/"), "/", cuffout)
    
## pull accession numbers off the files
geo_accession <- gsub( "_.*$", "", cuffout)

## read in the cuffout files
for (i in 1:length(cuff.files)) {
    assign(geo_accession[i], 
           read.table(cuff.files[i], 
                      header = TRUE)
    ) 
    }
```

```{r}
### manipulate supp files (already read in)

## cat these patients together also 
## filter out unknown gene_ids, select gene_ids and FPKM col from each dataframe (each df is a geo_accession/ patient)

## empty df for loop
expression.set <- data.frame(geo_accession = character(), gene_id = character(), locus = character(), FPKM= numeric())

## loop through the indiv. geo_access, pull out gene and FPKMs
for (i in 1:length(geo_accession)) {
    expression.set <- expression.set %>%
        bind_rows(., eval(as.symbol(geo_accession[i])) %>%
                      mutate(geo_accession = geo_accession[i]) %>%
                      select(geo_accession, gene_id, locus, FPKM) %>%
                      filter(!str_detect(gene_id, "CUFF.*")) 
        )
}

## pivot wider to match MeningiOMICS harmonized expression format
gse85133_exprs <- expression.set %>%
    mutate(tpm = exp(log(FPKM) - log(sum(FPKM)) + log(1e6))) %>%
    select(-FPKM) %>%
    pivot_wider(names_from = geo_accession, values_from = tpm) %>%
    select(-locus)
```

## Transform rnaseq Data
``` {r ExpressionDataTransform}
# function to find standard deviation of each probe across all tumors
rowsd <- function(x) {
                      sqrt(
                        rowSums((x - rowMeans(x, na.rm = TRUE))^2, na.rm = TRUE)/
                          (dim(x)[2] - 1)
                        )
}

gse85133_rnaseq <- gse85133_exprs %>% 
               rename("gene" = "gene_id") %>%                       # rename name variable as 'gene' 
               filter(gene != "") %>%                                         # keep only data for named genes
               mutate(sd = rowsd(.[,-1])) %>%                                 # calculate sd for each probe, index out gene name
               group_by(gene) %>%                                             # group probes by gene
               filter(sd == max(sd)) %>%                                      # keep the probe with the max variation for each gene
               filter(row_number() == 1) %>%
               ungroup() %>%                                                  # ungroup by genes 
               select(-sd) %>%                                                # select all but sd column
               pivot_longer(-gene,                                            # do not group by gene 
                            names_to = "geo_accession",                       
                            values_to = "tpm") %>%                          
               mutate(log_tpm = log2(tpm + 0.000001)) %>%                 # log2 transform  for variance stabilization 
               group_by(gene) %>%                                             # standardize for cross data set comparison
       #        mutate("standardized" = (log_tpm - mean(log_tpm, na.rm = T))/               # center at mean for gene 
       #                                  sd(log_tpm, na.rm = T),                          # scale by sd for gene
       #              ) %>%
                ungroup() %>%                                                 # ungroup by genes 
                select(-tpm) %>% #-log_tpm) %>%                                           # select all but sd column
                pivot_wider(names_from = "gene",                              # final format genes as rows
                            values_from = log_tpm) #"standardized")                     # final format expression across accession as columns    
```

## Dataset for Analysis
```{r FinalDataset}
## combine phenotype and expression data for workable data
gse85133_data <- inner_join(gse85133_pheno, gse85133_rnaseq, 
                            by = "geo_accession")   # match by accession

## create metatag for dataset
gse85133_info <-  
                list(
                     name = "Harmanci_a",               # author
                     accession = "GSE85133",         # dataset accession 
                     # the following are logical to determine clinical data available for dataset
                     geo_accession = "geo_accession" %in% colnames(gse85133_pheno),
                      sample_id = "sample_id" %in% colnames(gse85133_pheno),
                      age = "age" %in% colnames(gse85133_pheno),
                      sex = "sex" %in% colnames(gse85133_pheno),
                      grade = "grade" %in% colnames(gse85133_pheno),
                      histology = "histology" %in% colnames(gse85133_pheno),
                      type = "type" %in% colnames(gse85133_pheno),
                      location = "location" %in% colnames(gse85133_pheno),
                      mib1_index = "mib1_index" %in% colnames(gse85133_pheno),
                      recurrence = "recurrence" %in% colnames(gse85133_pheno),
                      recurrence_yrs = "recurrence_yrs" %in% colnames(gse85133_pheno),
                      recurrence_freq = "recurrence_freq" %in% colnames(gse85133_pheno),
                      followup_days = "followup_days" %in% colnames(gse85133_pheno),
                      survival_yrs = "survival_yrs" %in% colnames(gse85133_pheno),
                      vital_indicator = "vital_indicator" %in% colnames(gse85133_pheno),
                      tissue = "tissue" %in% colnames(gse85133_pheno),
                      progression = "progression" %in% colnames(gse85133_pheno),
                      radiation = "radiation" %in% colnames(gse85133_pheno),
                      rna_integrity ="rna_integrity" %in% colnames(gse85133_pheno),

                     sample_numbers = nrow(gse85133_data) # number of samples 
                     )

## create dataset as list 
Clark_V_2016_rnaseq <- list(
                dataset_info = gse85133_info,             # metatag
                clinical_data = gse85133_pheno,           # clinical data
                expression_data = gse85133_rnaseq,           # expression data
                data = gse85133_data #,                     # combined, workable data
             #   full_GEO = gse85133                       # full original download from GEO
)
```

```{r}
save(Clark_V_2016_rnaseq, file = here("datasets/rnaseq","Clark_V_2016_rnaseq.Rda")) 
```