---
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: false
---


```{r setup, include=FALSE}
## library
library(knitr)
library(DT)
library(htmlTable)
library(gtsummary)
library(flextable)
library(dplyr)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


```





## Directions for Use
        
All analyses are built to react to user selection and are dynamically generated based on user selection. In the analyses tabs, make selections for variables of interest in order from the top down. Start by selecting a gene or genes of interest. The application will find all of the clinical variables which exist in datasets which contain the gene(s) of interest. Next, select a dataset; the application will only display datasets which have the selected clinical variable. Users can change the selected gene or genes of interest at anytime; the available clincal variables and datasets will be recalculated for selection.
         
        
*****
        
## Future Features
        
Currently there are no datasets included in MeningiOMICS available for robust survival analysis. This feature will be added once sufficient data becomes available.

*****
        
## A Note on Harmonization and Collective Analysis

Great care was taken to preserve the integrity of each dataset included in MeningiOMICS. Variable names and levels of factor variables were altered for some datasets and expression values were log2 transformed during the harmonization stage. Expression values were also centered and scaled for each gene in each experiment to facilitate ease of collective analysis. Because the datasets originated from a variety of experimental designs and were processed on differing platforms, combined analysis is not recommended; as such, the tool does not allow 'pooled' or 'overall' analyses.

*****       
  
## Links to Data Sources

The table below contains links to the GEO Accession page for each dataset.

```{r}
## make a clicakable link
clickable <- function(link, text) {
    paste0("<a href='", link, "' target='_blank'>", text, "</a>")
}
```


```{r}
## array datasets
geo_link_array <- rbind(
      c("Clark_V_2013",    "GSE58037", "https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE58037"),
      c("Clark_V_2016",    "GSE85135", "https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE85135"),
      c("Schulten_H",      "GSE77259", "https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE77259"),
      c( "Schmidt_M",      "GSE74385", "https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE74385"),
      c( "TorresMartin_M", "GSE54934", "https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE54934"),
      c( "Yeh_T",          "GSE55609", "https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE55609"),
      c( "DUrso_O",        "GSE16156", "https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE16156"),
      c( "Lee_Y",          "GSE16581", "https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE16581"),
      c( "Keller_A",       "GSE12530", "https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE12530"),
      c( "Claus_E",        "GSE9438",  "https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE9438"),
      c( "Scheck_A",       "GSE4780",  "https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE4780")
      ) %>% 
    data.frame() %>% 
    setNames(c("dataset", "geo_accession", "geo_link"))  %>% 
    mutate(`Link to GEOAccession Viewer` = clickable(link = geo_link, text = geo_accession),
           datatype = "Microarray") 
```


```{r}
## methylation datasets
geo_link_methyl <- rbind(
    c("Harmanci_a_2013",  "GSE91376", "https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE91376"),
    c("Agnihotri_S_2013", "GSE83933", "https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE83933")
    ) %>% 
    data.frame() %>% 
    setNames(c("dataset", "geo_accession", "geo_link")) %>% 
    mutate(`Link to GEOAccession Viewer` = clickable(link = geo_link, text = geo_accession),
           datatype = "Methylation") 


```

```{r}
## combine all datatypes
master_set <- bind_rows(geo_link_array,
                        geo_link_methyl)
```

```{r}
## manipulate for presentation    
references <- master_set %>% 
    select(datatype, dataset, `Link to GEOAccession Viewer`) %>%     
    mutate(datatype = factor(datatype))  ## factor any cols for selectable buttons in datatable

## table of links
datatable(references, 
          escape = FALSE,  # allows the html clickables to render
          filter = "top",  # search boxes at top (allows filtering by datatype)
          rownames = FALSE, # no row names
          options = list(columnDefs = list(list(width = '200px', targets = "_all"))) # sets colwidths instead of auto b/c filtering issue
          ) 
```


*****

## Summary of Data from Data Sources
      
This table describes the data available the datasets included in MeningiOMICS.

```{r}
## load datasets to be summarized
array_datasets_load <- list.files(path = "datasets/array", full.names = TRUE)

#### load datasets
invisible(lapply(array_datasets_load, 
       load, 
       .GlobalEnv)
)

## load datasets to be summarized
methyl_datasets_load <- list.files(path = "datasets/methylation", full.names = TRUE)

#### load datasets
invisible(lapply(methyl_datasets_load, 
       load, 
       .GlobalEnv)
)
```


```{r}
# write a loop to to create gt_summary tables
summaries <- list()
titles <- vector(mode = "character", length = length(master_set$dataset))
for (i in 1:length(master_set$dataset)) {
    summaries[[i]] <- 
        eval(as.symbol(master_set$dataset[i]))$clinical_data %>% 
        select(!c(geo_accession,sample_id)) %>% 
        tbl_summary(missing_text = "(Missing)") %>% 
        modify_header(label = "**Clincal Variable**") %>%
        modify_footnote(all_stat_cols() ~ "n (%); Median (IQR)") %>% 
        bold_labels()
    titles[i] <- paste0(master_set$datatype[i],"\n",master_set$dataset[i],"\n", master_set$geo_accession[i])
}

tbl_merge(summaries, tab_spanner = titles) %>% 
    modify_spanning_header(list(label ~ "Datatype\nDataset\nGEO Accession")) %>% 
    as_flex_table() %>% 
    theme_booktabs() 



```
