---
title: "Standard object for dataset"
author: "Anh N Tran"
date: 9/20/2019
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Standard object for datasets

A standard dataset is a named list of 3: `metadata`, `pdata`, `value`

```{r dataset}
dataset <- list(metadata = NULL,
                pdata = NULL,
                value = NULL)
```


##. `metadata`
A named list of the following components (unavailable data will be a vector of 1 NA)
    - `dataset_name`: dataset name, character vector of 1
    - `geo_series`: GEO series, character vector of 1
    - `geo_subseries`: GEO subseries, character vector of 1
    - `geo_platform`: GEO platform, character vector of 1
    - `data_type`: types of data, character vector of 1
    - `value_type`: type/unit of value, character vector of 1
    - `sample_numbers`: number of samples, named numeric vector of 3 (`non_tumor`, `tumor`, `other`)
    - `clinical_data`: denote of certain clinical data is available, named logical vector of (`age`, `sex`, `grade`, `histology`, `recurrence`, `survival`)
    - `species`: Latin name for species
    
```{r metadata}
dataset$metadata <- list (dataset_name = "name",
                          geo_series = "GSExxxxxx",
                          geo_subseries = "GSExxxxxx",
                          geo_platform = "GPLxxxxxx",
                          data_type = "rna",
                          value_type = "Normalized signal intensity",
                          sample_numbers = c(non_tumor = 0,
                                             tumor = 0,
                                             other = 0),
                          clinical_data = c(age = FALSE,
                                            sex = FALSE,
                                            grade = FALSE,
                                            histology = FALSE,
                                            recurrence = FALSE,
                                            survival = FALSE),
                          species = "Homo sapiens")
```

## `pdata`
Chinical data. A data frame, a sample on each row. First column is sample names.

```{r pdata}
pdata_example <- data.frame(sample = c("sample_1", "sample_2", "sample_3"),
                            age = c(41, 42, 43),
                            grade = c("I", "II", "III"),
                            recurrence = c("primary", "secondary", NA))

dataset$pdata <- pdata_example
```


## `value`
Value of expression of each gene for each sample. A data frame, a sample on each row. First column is sample names.

```{r value}
value_example <- data.frame(sample = c("sample_1", "sample_2", "sample_3"),
                            GENE1 = c(1, 2, 3),
                            GENE2 = c(2, NA, 3),
                            GENE3 = c(1, 1, 2))

dataset$value <- value_example
```

```{r}
summary(dataset)

#          Length Class      Mode
# metadata 9      -none-     list
# pdata    4      data.frame list
# value    4      data.frame list

str(dataset)

# List of 3
#  $ metadata:List of 9
#   ..$ dataset_name  : chr "name"
#   ..$ geo_series    : chr "GSExxxxxx"
#   ..$ geo_subseries : chr "GSExxxxxx"
#   ..$ geo_platform  : chr "GPLxxxxxx"
#   ..$ data_type     : chr "rna"
#   ..$ value_type    : chr "Normalized signal intensity"
#   ..$ sample_numbers: Named num [1:3] 0 0 0
#   .. ..- attr(*, "names")= chr [1:3] "non_tumor" "tumor" "other"
#   ..$ clinical_data : Named logi [1:6] FALSE FALSE FALSE FALSE FALSE FALSE
#   .. ..- attr(*, "names")= chr [1:6] "age" "sex" "grade" "histology" ...
#   ..$ species       : chr "Homo sapiens"
#  $ pdata   :'data.frame':	3 obs. of  4 variables:
#   ..$ sample    : Factor w/ 3 levels "sample_1","sample_2",..: 1 2 3
#   ..$ age       : num [1:3] 41 42 43
#   ..$ grade     : Factor w/ 3 levels "I","II","III": 1 2 3
#   ..$ recurrence: Factor w/ 2 levels "primary","secondary": 1 2 NA
#  $ value   :'data.frame':	3 obs. of  4 variables:
#   ..$ sample: Factor w/ 3 levels "sample_1","sample_2",..: 1 2 3
#   ..$ GENE1 : num [1:3] 1 2 3
#   ..$ GENE2 : num [1:3] 2 NA 3
#   ..$ GENE3 : num [1:3] 1 1 2
```

