---
title: "Anova across grades"
author: "Kaitlyn O'Shea"
date: "12/2/2019"
output: html_document
---

### Libraries needed
```{r}
library(ggplot2)
library(dplyr)
library(car)
library(stringr)
```

## Load Datasets
```{r}
# load("~/Documents/Thesis/mgm_eda/datasets/GSE85135_Clark_V_KO.Rda")
# load("~/Documents/Thesis/mgm_eda/datasets/GSE58037_Clark_V_KO.Rda")
# load("~/Documents/Thesis/mgm_eda/datasets/GSE77259_Schulten_H_KO.Rda")
load("~/Documents/Thesis/mgm_eda/datasets/GSE74385_Schmidt_M_KO.Rda")
# load("~/Documents/Thesis/mgm_eda/datasets/GSE54934_Torres-Martin_M_KO.Rda")
# load("~/Documents/Thesis/mgm_eda/datasets/GSE55609_Yeh_T_KO.Rda")
# load("~/Documents/Thesis/mgm_eda/datasets/GSE16156_DUrso_O_KO.Rda")
# load("~/Documents/Thesis/mgm_eda/datasets/GSE16581_Lee_Y_KO.Rda")
# load("~/Documents/Thesis/mgm_eda/datasets/GSE12530_Keller_A_KO.Rda")
# load("~/Documents/Thesis/mgm_eda/datasets/GSE9438_Claus_E_KO.Rda")
# load("~/Documents/Thesis/mgm_eda/datasets/GSE4780_Scheck_A_KO.Rda")
```

## Proof of concept using GSE74385_Schmidt_M_KO.Rda

### Visualize expression
```{r}
## access data
schm <- Schmidt_M[["data"]]

## Pick a gene - use PTTG1 as literature suggests association with more agressive (higher grade) 
# schm$PTTG1 # great! it exists

## Summary statistics across grade
summ <- schm %>% 
                group_by(grade) %>% 
                                    summarize(mean = mean(PTTG1), 
                                              median = median(PTTG1), 
                                              sd = sd(PTTG1),
                                              total = n()
                                              )
summ

## Visualize expression values across grade
ggplot(data = schm, 
       aes(x = grade, 
           y = PTTG1)) +
    geom_boxplot() +
    theme_bw() +
    ggtitle("Expression of PTTG1 by WHO Grade") +
    xlab("WHO Grade") +
    ylab("Expression") +
    stat_summary(fun.y=mean,                      # add in a blue dashed line at the mean
                 geom = "errorbar", 
                 aes(ymax = ..y.., ymin = ..y..),
                 width = .75, 
                 linetype = "dashed",
                 color = "blue")

cbind(summ, d) # doesn't work because pairwise is joined with single grade - duh
```
*report summary and pairwaise and overall aov*

## Determine Significance of Difference (Anova)

### Start with familiar approach
```{r}
fit <- lm(data = schm,    # linear model  
          PTTG1 ~ grade)  # expression over grade
summary(fit)   
```

```{r}
plot(fit, which = 1)      # check for equal variance

leveneTest(data = schm,   # check for homogeneity of variance
           PTTG1 ~ grade
           )
```

```{r}
anova(fit) # base r

Anova(fit) # car package
```
### Do this with fewer steps
```{r}
summary(aov(data = schm,     
          PTTG1 ~ grade))
```
### Check pairwise significance
```{r}
TukeyHSD(aov(data = schm,     
          PTTG1 ~ grade))
```

## Gather information for presentation - incomplete
```{r}
a <- summary(aov(data = schm,     
          PTTG1 ~ grade)
        )

b <- data.frame(a[[1]]$'Pr(>F)'[1])
names(b) <- c("p adj")
row.names(b) <- "overall"
b # doesn't need to be on the graph
```

```{r}
c <- TukeyHSD(aov(data = schm,     
          PTTG1 ~ grade))
str(c)

d <- as.data.frame(c$grade)
d
library(plyr)
e <- rbind.fill(d, b)
e
```

## Function to do this for any gene
```{r}
var_grade <- function(gene, dataset) {
    print(gene)
    print(as.formula(paste0(gene,"~grade")))
                                      aov <- summary(aov(data = dataset,                        # aov, call the dataset
                                                         formula = as.formula(paste0(gene,      # formula to take gene input
                                                                                    "~grade"))  # interpret grade as variable
                                                         )
                                                     )
                                      hsd <- TukeyHSD(aov(data = dataset,                       # hsd, call the dataset
                                                         formula = as.formula(paste0(gene,      # formula to take gene input
                                                                                    "~grade"))  # interpret grade as variable
                                                         )
                                                     )
                                      print(aov)                                                # print aov 
                                      print(hsd)                                                # print hsd
                                      }
```

### Proof of concept's concept
```{r}
var_grade(gene = "PTTG1", dataset = schm)
```

## Now do it for more than one gene
```{r eval=FALSE, include=FALSE}
R_HSA_210746 <- "NEUROD1;PAX4;INSM1;NKX-2;NEUROG3"
split_R_HSA_210746 <- str_split(R_HSA_210746,
                                ";",
                                simplify = T)

genes <- sapply(X = split_R_HSA_210746,
       FUN = function(X) {
                          gene = X
                          print(gene)
                        }
)

sapply(genes, var_grade, dataset= schm)
```

