---
title: "Anova across grades"
author: "Kaitlyn O'Shea"
date: "12/2/2019"
output: html_document
---

### Libraries needed
```{r}
library(ggplot2)
library(dplyr)
library(car)
library(stringr)
```

## Load Datasets
```{r}
# load("datasets/GSE85135_Clark_V_KO.Rda")
# load("datasets/GSE58037_Clark_V_KO.Rda")
# load("datasets/GSE77259_Schulten_H_KO.Rda")
load("datasets/GSE74385_Schmidt_M_KO.Rda")
# load("datasets/GSE54934_Torres-Martin_M_KO.Rda")
# load("datasets/GSE55609_Yeh_T_KO.Rda")
# load("datasets/GSE16156_DUrso_O_KO.Rda")
load("datasets/GSE16581_Lee_Y_KO.Rda")
# load("datasets/GSE12530_Keller_A_KO.Rda")
# load("datasets/GSE9438_Claus_E_KO.Rda")
# load("datasets/GSE4780_Scheck_A_KO.Rda")
```

## Proof of concept using GSE74385_Schmidt_M_KO.Rda

### Visualize expression
```{r}
## access data
schm <- Schmidt_M[["data"]]

## Pick a gene - use PTTG1 as literature suggests association with more agressive (higher grade) 
# schm$PTTG1 # great! it exists

## Summary statistics across grade
summ <- schm %>% 
                group_by(grade) %>% 
                                    summarize(mean = mean(PTTG1), 
                                              median = median(PTTG1), 
                                              sd = sd(PTTG1),
                                              total = n()
                                              )
summ

## Visualize expression values across grade
ggplot(data = schm, 
       aes(x = grade, 
           y = PTTG1)) +
    geom_boxplot() +
    theme_bw() +
    ggtitle("Expression of PTTG1 by WHO Grade") +
    xlab("WHO Grade") +
    ylab("Expression") +
    stat_summary(fun.y=mean,                      # add in a blue dashed line at the mean
                 geom = "errorbar", 
                 aes(ymax = ..y.., ymin = ..y..),
                 width = .75, 
                 linetype = "dashed",
                 color = "blue")

#cbind(summ, d) # doesn't work because pairwise is joined with single grade - duh
```
*report summary and pairwaise and overall aov*

## Determine Significance of Difference (Anova)

### Start with familiar approach
```{r}
fit <- lm(data = schm,    # linear model  
          PTTG1 ~ grade)  # expression over grade
summary(fit)   
```

```{r}
plot(fit, which = 1)      # check for equal variance

leveneTest(data = schm,   # check for homogeneity of variance
           PTTG1 ~ grade
           )

leveneTest(aov(data = schm, ## leveneTest (Brown-Forsythe - centered at median) takes an aov object - already in the app! woot    
          PTTG1 ~ grade))
```
```{r}
x <- leveneTest(aov(data = schm,    
          PTTG1 ~ grade)
)
names(x)

y <- x$`Pr(>F)`[[1]]
y
```



```{r}
anova(fit) # base r

Anova(fit) # car package
```
### Do this with fewer steps
```{r}
summary(aov(data = schm,     
          PTTG1 ~ grade))
```
### Check pairwise significance
```{r}
TukeyHSD(aov(data = schm,     
          PTTG1 ~ grade))
```

## Gather information for presentation - incomplete
```{r}
a <- summary(aov(data = schm,     
          PTTG1 ~ grade)
        )

b <- data.frame(a[[1]]$'Pr(>F)'[1])
names(b) <- c("p adj")
row.names(b) <- "overall"
b # doesn't need to be on the graph
```

```{r}
c <- TukeyHSD(aov(data = schm,     
          PTTG1 ~ grade))
str(c)

d <- as.data.frame(c$grade)
d[,c(1,4)]

library(plyr)
e <- rbind.fill(d, b)
e
```

## Function to do this for any gene
```{r}
var_grade <- function(gene, dataset) {
    print(gene)
    print(as.formula(paste0(gene,"~grade")))
                                      aov <- summary(aov(data = dataset,                        # aov, call the dataset
                                                         formula = as.formula(paste0(gene,      # formula to take gene input
                                                                                    "~grade"))  # interpret grade as variable
                                                         )
                                                     )
                                      hsd <- TukeyHSD(aov(data = dataset,                       # hsd, call the dataset
                                                         formula = as.formula(paste0(gene,      # formula to take gene input
                                                                                    "~grade"))  # interpret grade as variable
                                                         )
                                                     )
                                      print(aov)                                                # print aov 
                                      print(hsd)                                                # print hsd
                                      }
```

### Proof of concept's concept
```{r}
var_grade(gene = "PTTG1", dataset = schm)
```

## Now do it for more than one gene
```{r eval=FALSE, include=FALSE}
R_HSA_210746 <- "NEUROD1;PAX4;INSM1;NKX-2;NEUROG3"
split_R_HSA_210746 <- str_split(R_HSA_210746,
                                ";",
                                simplify = T)

R_HSA_418555 <- str_replace_all("OR7G1;OR9K2;OR2M7;OR2M5;OR2M4;OR2M3;OR2M2;OR52N2;OR5D18;OR10W1;OR5D16;OR5B3;OR5B2;OR5D14;OR5D13;OR8H3;OR8H2;OR8H1;OR2L8;OR1J2;OR2L3;TAAR5;OR2L2;OR5R1;OR6T1;OR10V1;OR4C12;OR6C6;OR6C4;OR8G5;OR6C3;OR4C16;OR6C2;OR4C15;OR5AK2;OR9I1;OR2K2;OR10T2;CRHR2;OR5M11;OR5M10;OR7D4;GPR15;OR4N5;OR2J3;OR2J2;OR4N4;OR10S1;OR4N2;OR52K2;OR9G9;OR4D11;OR10C1;OR9G4;OR9G1;GPR27;OR10J1;OR11L1;GPR25;OR10J3;OR10J5;OR5W2;OR6Y1;OR5AP2;OR5B12;OR52B2;OR5B17;OR4D5;OR4D1;OR1N1;OR10Z1;OR6X1;OR52R1;OR4C6;OR3A4;OR3A3;OR4A15;OR4A16;OR8K5;OR8K3;OR10H1;OR8K1;OR10H2;OR10H5;ADCY8;OR1M1;OR4S2;ADCY5;NPS;GPR150;OR4C45;OR5AN1;GPHA2;OR1F2P;OR4C46;OR6F1;OR8J3;OR8J1;OR14A16;OR10G4;OR51M1;OR1L3;OR5T3;OR5T2;OR5T1;OR10X1;ADM2;MC5R;OR10G8;OR4A5;OR10G9;OR1C1;OR5K4;OR5K2;OR5K1;OR6M1;OR2T12;OR2T10;OR5H15;OR1S2;OR1S1;OR13D1;OR1B1;OR2D2;OR14C36;OR5J2;OR5AS1;OR52E2;OR2T8;OR2T6;OR2T5;OR2T4;OR2T3;OR13C8;OR4X2;OR2L13;OR5AC2;OR51T1;OR13C3;OR1A2;OR1A1;OR6K6;OR6K3;OR5AR1;OR6K2;OR5I1;GPR84;OR1Q1;OR2AK2;OR4A47;OR4F6;OR5H6;DRD5;OR10K2;OR10K1;OR5H1;RLN2;OR2A7;OR51A2;RLN1;OR51A7;OR14J1;OR8U8;OR1G1;OR4M2;OR4M1;OXT;OR10AG1;OR8U1;OR8D4;OR8D1;TSHB;OR51G1;OR6C75;OR2H2;OR1F1;OR6C74;OR13G1;OR2H1;OR14I1;OR6P1;GIP;OR2B11;OR10A3;OR10A5;OR2G6;OR5M9;OR51F1;OR5M8;OR2G3;OR1E2;OR6C65;OR2G2;OR1E1;OR5M3;OR5M1;OR10P1;OR2T34;OR2T33;OR2W3;OR2W1;OR8B4;GNG8;OR8B3;OR6C68;OR13F1;OR12D2;OR1D5;OR2T27;OR1D4;OR2T29;OR5L2;OR6N2;OR5L1;OR6N1;OR51V1;OR11A1",
                          ";",
                          ",")

cat(paste(R_HSA_418555))

genes <- sapply(X = split_R_HSA_210746,
       FUN = function(X) {
                          gene = X
                          print(gene)
                        }
)

sapply(genes, var_grade, dataset= schm)
```

### welch one-way test
```{r}
oneway.test(data = schm,     
          PTTG1 ~ grade)

v <- aov(data = schm,PTTG1 ~ grade)

#names(v)
w <- v$call
#w
#str(w)

r <- as.character(v$call)
#r
#str(r)
e <- r[[2]]
f <- r[[3]]

t <- oneway.test(formula = as.formula(e), data = eval(as.symbol(f)))
names(t)

str(t$p.value)

```



```{r}
is.numeric(schm %>% select(grade))

is.double(schm %>% select(grade))

```
```{r}
lee <- Lee_Y[["data"]]

## Summary statistics across grade
summ2 <- lee %>% 
               summarize(
                   mean = mean(PTTG1), 
                   median = median(PTTG1), 
                   sd = sd(PTTG1),
                   total = n()
               )
summ2

test <- summary(aov(data = lee, PTTG1 ~ grade))

test # pvalue small
```

```{r}
fit_cont <- summary(lm(data = lee,  # linear model  
                       PTTG1 ~ age) # expression over age
) 

lm <- data.frame(fit_cont$coefficients)[2,c(1,4)]
colnames(lm) <- c("V1","V2")
```

## create spearman correlation
```{r}
t <- cor.test(lee$PTTG1, lee$age, method = "spearman")

rho <- as.data.frame(cbind(t$estimate[[1]], t$p.value))

cont <- rbind(lm,rho)
names(cont) <- c("Estimate", "p-value")
row.names(cont) <- c("Univariable Linear Slope", "Correlation Coefficient")
cont
```
```{r}
str(cont)
is.numeric(lee$age)
```
```{r}
lee$PTTG1
lee$age
```

